---
title: "Data Preparation for Maxent Models"
author: "Katharine Egan"
email: "katharineegan34@gmail.com"
date: "July 2018"
output: html_document
---

This script prepares the data for modeling input. 

Steps: 
1) Read in and stack the predictor data (rasters)
2) Remove landslide presence points that fall outside of the country shapefile of Myanmar
3) Extract presence points from the raster stack - each point will have a value from the raster stack associated with it
4) Randomly select 10,000 background points from the raster stack (these are considered the "absences" in a Maxent model) - after this, each point will already have a value associated with it from the raster stack
5) Save presence points and background points to CSVs for use in the next script.

Define variables here: 
```{r}
# define your raster temporary folder; this is where the script will store temporary raster files while raster things are happening
raster_temp_folder <- "E:/SERVIR/Landslide_Models/temp"

# define folder where the ascii files are located
ascii_path <- "E:/SERVIR/Landslide_Models/predictors_ascii"

# path to where the country shapefile is located 
country_shp_path <- "E:/SERVIR/Landslide_Models/layers"

# define name of the country shapefile 
country_shp_name <- "temporary_myanmar"

# path to landslide occurrences in a CSV format
# this file MUST have columns that look like this and in this order:
# Group, Longitude, Latitude
landslides_csv <- "E:/SERVIR/Landslide_Models/points/landslide_presences.csv"

# specify the path to save all RData files 
rdata_path <- "E:/SERVIR/Landslide_Models/models/Maxent/rdata_files"

# path to save background points and presence points as CSVs
points_path <- "E:/SERVIR/Landslide_Models/models/Maxent/points"

# vector of column names
# these MUST be the same exact names as the predictor ascii files 
column_names <- c("dem_tmp", "slope_tmp", "land_cover_tmp")
```

Libraries:
```{r}
library(raster)
library(rgeos)
library(rgdal)
library(sp)
library(dismo)
library(tidyverse)
library(readxl)
library(writexl)
library(MaxentVariableSelection)
library(beepr)
```

Define raster temporary folder here: 
```{r}
# set raster options 
rasterOptions(tmpdir = raster_temp_folder, progress = "text")
```

Step 1) Read in the predictors, landslide occurrences, and the country shapefile
```{r}
# read in the raster data 
raster_list <- list.files(path=ascii_path, 
     pattern =".asc$", full.names=TRUE)

# stack the rasters
rs <- stack(raster_list)

# set projection for the raster stack 
crs(rs) <- "+proj=utm +zone=47 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

# read in model extent for trimming points
model_fp <- readOGR(dsn = country_shp_path, layer = country_shp_name)

# read in the occurrence data 
landslides <- read_csv(landslides_csv)
```

Step 2) This chunk of code removes data points that fall outside of the country shapefile and also removes any duplicate coordinates
```{r}
# make a copy of the occurrence data
landslides_sp <- landslides

# convert to a spatial object
coordinates(landslides_sp) <- ~Longitude+Latitude

# it MUST have the same projection as the country shapefile and raster stack
proj4string(landslides_sp) <- "+proj=utm +zone=47 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

# look for points that fall outside country shapefile
# where there are NAs, those points are inside the model extent
tmp <- over(landslides_sp, model_fp)
tmp <- tmp %>% tibble::rownames_to_column() %>% 
  mutate_if(is.factor, as.numeric)

# combine the original data with the one produced by over()
landslides_trim <- landslides %>% tibble::rownames_to_column() %>% 
  full_join(tmp) %>% 
  filter(!is.na(Id)) %>% 
  #filter(!is.na(LON))
  select(Group, Longitude, Latitude)

# convert to a spatial object and check to make sure points are inside the country shapefile
tmp_sp <- landslides_trim
coordinates(tmp_sp) <- ~Longitude+Latitude
proj4string(tmp_sp) <- "+proj=utm +zone=47 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

# create two separate data frames: one for joining and one for extrating from the raster stack 
ls_join <- landslides_trim %>% distinct() %>% unique %>% 
  tibble::rownames_to_column(var = "ID")
landslides_trim <- landslides_trim %>% distinct() %>% unique %>% select(-Group) 
```

Step 3) Format data for MaxentVariableSelect package; need two separate CSVs for input into the MaxentVariableSelect package: an occurrence one and a background point one
```{r}
# extract presence locations from raster stack
# each presence should be associated with a value from the raster
landslides_trim_env <- raster::extract(rs, landslides_trim)

# clean up the extracted data frame
# first three columns must be "species, longitude, latitude"
# need to join back with the original data
# HAVE TO REMOVE NAs - won't work with the maxent R package if there are NAs
landslide_trim_env_clean <- as.data.frame(landslides_trim_env) %>% 
  tibble::rownames_to_column(var = "ID") %>% 
  full_join(ls_join) %>% 
  rename(species = Group,
         longitude = Longitude,
         latitude = Latitude) %>% 
  drop_na()

# specify other columns 
other <- c("species", "longitude", "latitude")

# add together with the column names 
t <- c(other, column_names)

# rearrange columns  
landslide_trim_env_clean <-
  landslide_trim_env_clean[,dput(as.character(t))]
```

Step 4) Manually select random background points
```{r}
# select background points from a raster object; the number provided in the set.seed() function determines the random sample, using the same number will produce the same random sample		

# use this code before the sampleRandom function every time, if you want to get the same "random samples"
# select 12,000 background points
set.seed(1) 
bg <- sampleRandom(x=rs,
                   size=10000,
                   na.rm=T, #removes the 'Not Applicable' points  
                   sp=T) # return spatial points 

# convert the background points to a dataframe
# clean this up, add in the original data 
# first three columns must be "species, longitude, latitude"
# need to join back with the original data
# HAVE TO REMOVE NAs - won't work with the maxent R package if there are NAs
a_clean <- bg@data %>% bind_cols(as.data.frame(bg@coords)) %>% 
  rename(longitude = x, latitude = y) %>% 
  drop_na() %>% 
  mutate(species = "bg") %>% 
  drop_na()

# specify other columns 
other <- c("species", "longitude", "latitude")

# add together with the column names 
d <- c(other, column_names)

# rearrange columns  
a_clean <- a_clean[,dput(as.character(d))]
```

Step 5) Save everything:
```{r}
# save the entire workspace
save(list = ls(all.names = TRUE), 
     file = paste(rdata_path, "/1_data_prep_workspace.RData", sep = ""),
     envir = .GlobalEnv)

# save raster stack as an rdata file
save(rs, file = paste(rdata_path, "/raster_stack_of_predictors.RData", 
                      sep = ""))

# write landslide occurrences to a csv
write_csv(landslide_trim_env_clean, 
          paste(points_path, "/occurrences.csv", sep = ""))

# write background points to a csv
write_csv(a_clean, paste(points_path, "/backgroundpoints.csv", sep = ""))

# play a noise to indicate when the entire script is done running
beep(3)
```

