---
title: "Final Maxent Model"
author: "Katharine Egan"
email: "katharineegan34@gmail.com"
date: "July 2018"
output: html_document
---

-Read in the predictors that compose the best model according to Script #2
-Stack predictors
-Read in presence data and background points from Script #1 
-Extract p/bg points from raster stack
-Using “dismo” package to run Maxent model in R
-Downloads a maxent.jar file from Maxent github and moves it to the dismo R package folder
-Dismo accesses Maxent and allows you to manipulate the settings in R
-Run model - USES THE FULL DATASET (no training/testing, get that data from Script #2)
-Only 1 model run (no replicates) for this script, but can specify replicates
-Use model to predict over the raster stack using raster::predict()
Makes logistic prediction raster - writes out a geotiff

Define variables here: 
```{r}
# define your raster temporary folder; this is where the script will store temporary raster files while raster things are happening
raster_temp_folder <- "E:/SERVIR/Landslide_Models/temp"

# read in the background points and the presence data 
p_path <- "E:/SERVIR/Landslide_Models/models/Maxent/points/occurrences.csv"
bg_path <- "E:/SERVIR/Landslide_Models/models/Maxent/points/backgroundpoints.csv" 

# path where all the data will be written out 
output <- "E:/SERVIR/Landslide_Models/models/Maxent/output/3_final_maxent_model_output"

# specify categorical data here 
# must be the same name 
categorical <- c("land_cover_tmp")
```

Libraries:
```{r}
library(raster)
library(rgeos)
library(rgdal)
library(sp)
library(dismo)
library(tidyverse)
library(readxl)
library(writexl)
library(MaxentVariableSelection)
library(beepr)
```

Define raster temporary folder here: 
```{r}
# set raster options 
rasterOptions(tmpdir = raster_temp_folder, progress = "text")
```

Stacking rasters for the model:
```{r}
# define the rasters that compose of the best model 
var1_path <- "E:/SERVIR/Landslide_Models/predictors_ascii/land_cover_tmp.asc"
var2_path <- "E:/SERVIR/Landslide_Models/predictors_ascii/dem_tmp.asc"
var3_path <- "E:/SERVIR/Landslide_Models/predictors_ascii/slope_tmp.asc"

# define as rasters
land_cover <- raster(var1_path)
dem <- raster(var2_path)
slope <- raster(var3_path)
# .... add more here depending on model

# stack rasters 
rs <- stack(land_cover, dem, slope)
#....add more variable names here depending on model

# set projection for the raster stack 
crs(rs) <- "+proj=utm +zone=47 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

Combine presence and background point data and subset out environmental data: 
```{r}
# read in the presence and absence data
p <- read_csv(p_path)
bg <- read_csv(bg_path)

# combine p and a
p_bg <- bind_rows(p, bg) %>%
  mutate(species = if_else(species == "landslide", 1, 0)) %>% 
  dplyr::rename(Group = species, 
                Longitude = longitude, 
                Latitude = latitude)

# subset out env data
env <- p_bg %>% 
  select(-c(Group, Longitude, Latitude))

# keep only landslide, longitude, and latitude in the pg file 
p_bg <- p_bg %>% 
  select(Group)
p_bg <- p_bg$Group
```

Run Maxent model: 
```{r}
# FIRST make sure you have rJava installed - if you are working on 64bit verison of R, you MUST have 64bit Java installed

# SECOND move your maxent jar file to the DISMO package folder
if(!file.exists(paste0(system.file("java",
                                    package="dismo"),"/maxent.jar"))  )   {
utils::download.file(url="https://raw.githubusercontent.com/mrmaxent/Maxent/master/ArchivedReleases/3.3.3k/maxent.jar",
                     destfile=paste0(system.file("java",
                                          package="dismo"),"/maxent.jar"),
                     mode="wb") ## wb for binary file, otherwise maxent.jar can not execute
}

# run maxent model with full dataset
mod <- dismo::maxent(x = env, ## env conditions
              p = p_bg,   ## 1:presence or 0:absence
              path = output,
              factors = categorical,
              removeDuplicates = TRUE,
              args = c("responsecurves=true", 
                       "removeduplicates=true",
                       "jackknife=true",
                       "outputformat=logistic",
                       "betamultiplier=1.0",
                       "maximumbackground=10000",
                       "replicates=1",
                       "replicatetype=subsample",
                       "writeplotdata=true"))
              
# view the maxent model in a html brower
mod
```

Predict over a raster stack, the maxent model in R does not write out an logistic ascii file and save to output location:
```{r}
# project to study area [raster]
pred <- predict(mod, rs) # studyArea is the clipped rasters 
plot(pred)

# save as a raster to the maxent_ouputs folder 
writeRaster(pred, 
            paste(output, "/landslide_logistic_predict.tif", sep=""),
            format = "GTiff")

# make a noise when done
beep(3)
```
