---
title: "Data preparation for logistic regression models"
author: "Katharine Egan"
email: "katharineegan34@gmail.com"
date: "July 2018"
output: html_document
---

-Read in presence and absence data 
-Read in the predictors that were used in Script #3 of the Maxent models and stack them
-Want to compare models - have to conduct the same model with same predictors
-Combine presence/absence data and extract from raster stack


Define variables here: 
```{r}
# define your raster temporary folder; this is where the script will store temporary raster files while raster things are happening
raster_temp_folder <- "E:/SERVIR/Landslide_Models/temp"

# read in the background points and the presence data 
p_path <- "E:/SERVIR/Landslide_Models/points/landslide_presences.csv"
a_path <- "E:/SERVIR/Landslide_Models/models/LR/points/landslide_absences.csv"
# specify the path to save all RData files 
rdata_path <- "E:/SERVIR/Landslide_Models/models/LR/rdata_files"

# path to save points to
points_path <- "E:/SERVIR/Landslide_Models/models/LR/points"

# define categorical variable here: 
# must match name of the raster
categorical_data <- c("land_cover_tmp")
```

Libraries you need: 
```{r}
library(raster)
library(sp)
library(rgdal)
library(tidyverse)
library(broom) # for cleaning up statistical objects
library(readxl)
library(writexl)
```

Define raster temporary folder here: 
```{r}
# set raster options 
rasterOptions(tmpdir = raster_temp_folder, progress = "text")
```

Stacking rasters for the model:
```{r}
# define the rasters that compose of the best model 
var1_path <- "E:/SERVIR/Landslide_Models/predictors_ascii/land_cover_tmp.asc"
var2_path <- "E:/SERVIR/Landslide_Models/predictors_ascii/dem_tmp.asc"
var3_path <- "E:/SERVIR/Landslide_Models/predictors_ascii/slope_tmp.asc"

# define as rasters
land_cover <- raster(var1_path)
dem <- raster(var2_path)
slope <- raster(var3_path)
# .... add more here depending on model

# stack rasters 
rs <- stack(land_cover, dem, slope)
#....add more variable names here depending on model

# set projection for the raster stack 
crs(rs) <- "+proj=utm +zone=47 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

Read in the data and clean up:
```{r}
# load presence and absence data
p <- read_csv(p_path)
a <- read_csv(a_path) 

# combine presence and absence data for extracting from raster stack
# add an ID to join with extracted df later on
pa <- bind_rows(p, a) %>% 
  tibble::rownames_to_column(var = "ID") %>% 
  mutate(ID = as.numeric(ID))

# subset out the lat/long only for extracting 
coords <- pa %>% dplyr::select(-c(ID,Group))
```

Extract point data from the raster stack:
```{r}
# extract points from the raster stack
# df = TRUE means output is a data frame
extract_coords_rs <- raster::extract(rs, coords, df = TRUE)

# join back with the presence/absence data
# remove any NAs
pa_rs <- extract_coords_rs %>% 
  full_join(pa) %>% 
  dplyr::select(-ID) %>% 
  mutate(Group = if_else(Group == "landslide", 1, 0)) %>%
  drop_na() 

# make sure categorical variables are factors/characters and NOT numeric
for (i in categorical_data){
  #print(i)
  pa_rs[, c(i)] <- as.factor(pa_rs[, c(i)])
}
```

Save everything
```{r}
# save the entire workspace
save(list = ls(all.names = TRUE), 
     file = paste(rdata_path, "/1_data_prep_workspace.RData", sep = ""),
     envir = .GlobalEnv)

# save raster stack as an rda file 
save(rs,
     file = paste(rdata_path, "/rs.RData", sep = ""))

# write out csv and excel of the point and predictor data as well
write_csv(pa_rs, paste(points_path, "/landslide_occurrences_with_env.csv",
                       sep = ""))
```
