---
title: "Area of logistic prediction output rasters for LR and Maxent"
author: "Katharine Egan"
email: "katharineegan34@gmail.com"
date: "July 2018"
output: html_document
---

Script to calculate the area of habitat suitability/probability of occurrence at certain thresholds 
0 - 0.25 | 0.25 - 0.5 | 0.5 - 0.75 | 0.75 - 1.0
Output: a table comparing area for each model as a CSV


Define variables here: 
```{r}
# raster temporary folder file path
raster_temp_folder <- "E:/SERVIR/Landslide_Models/temp"

# file path to maxent logistic prediction raster 
maxent_raster <- "E:/SERVIR/Landslide_Models/models/Maxent/output/3_final_maxent_model_output/landslide_logistic_predict.tif"

# file path to LR logistic prediction raster
LR_raster <- "E:/SERVIR/Landslide_Models/models/LR/output/logistic_prediction.tif"

# file path to write out the suitable habitat table 
table_path <- "E:/SERVIR/Landslide_Models/area_calculations"
```

Libraries:
```{r}
library(raster)
library(rgdal)
library(sp)
library(tidyverse)
library(rgeos)
```

Define raster temporary folder here: 
```{r}
# set raster options 
rasterOptions(tmpdir = raster_temp_folder, progress = "text")
```

Read in the logistic prediction rasters:
```{r}
# read in prediction rasters
mlr <- raster(LR_raster)
mx <- raster(maxent_raster)
```

Calculate area of suitable habitat or probability of landslide occurrences for certain thresholds: (this will take a while to run)
```{r}
# 0 to  0.25 
# reclassify cells first 
# matrix of values to be reclassified to
m <- c(0, 0.25, 1, 0.25, 1.00, 0)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

# reclassify the raster 
mlr1_reclass <- reclassify(mlr, rclmat)
mx1_reclass <- reclassify(mx, rclmat)

# sum number of pixels above 0
# multiply by the square of the resolution to get the area
mlr1_km <- (sum(mlr1_reclass[] > 0, na.rm = T) * res(mlr1_reclass)[1]^2) / 1000 
mx1_lm <- (sum(mx1_reclass[] > 0, na.rm = T) * res(mx1_reclass)[1]^2) / 1000

# get the percentage of cells
# number of cells above 0, divded by the total number of cells in the entire raster multipled by 100
mlr1_pc <- (sum(mlr1_reclass[] > 0, na.rm = T) / 238627168) * 100
mx1_pc <- (sum(mx1_reclass[] > 0, na.rm = T) / 238627168) * 100

# 0.25 to 0.5 
# reclassify cells first 
# matrix of values to be reclassified to
m <- c(0, 0.25, 0, 0.25, 0.5, 1, 0.5, 1.0, 0)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

# reclassify the raster 
mlr2_reclass <- reclassify(mlr, rclmat)
mx2_reclass <- reclassify(mx, rclmat)

# sum number of pixels above 0
# multiply by the square of the resolution to get the area
mlr2_km <- (sum(mlr2_reclass[] > 0, na.rm = T) * res(mlr2_reclass)[1]^2) / 1000 
mx2_km <- (sum(mx2_reclass[] > 0, na.rm = T) * res(mx2_reclass)[1]^2) / 1000

# get the percentage of cells
# number of cells above 0, divded by the total number of cells in the entire raster multipled by 100
mlr2_pc <- (sum(mlr2_reclass[] > 0, na.rm = T) / 238627168) * 100
mx2_pc <- (sum(mx2_reclass[] > 0, na.rm = T) / 238627168) * 100


# 0.5 to 0.75 
# reclassify cells first 
# matrix of values to be reclassified to
m <- c(0, 0.5, 0, 0.5, 0.75, 1, 0.75, 1.0, 0)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
rclmat

# reclassify the raster 
mlr3_reclass <- reclassify(mlr, rclmat)
mx3_reclass <- reclassify(mx, rclmat)

# sum number of pixels above 0
# multiply by the square of the resolution to get the area
mlr3_km <- (sum(mlr3_reclass[] > 0, na.rm = T) * res(mlr3_reclass)[1]^2) / 1000 
mx3_km <- (sum(mx3_reclass[] > 0, na.rm = T) * res(mx3_reclass)[1]^2) / 1000

# get the percentage of cells
# number of cells above 0, divded by the total number of cells in the entire raster multipled by 100
mlr3_pc <- (sum(mlr3_reclass[] > 0, na.rm = T) / 238627168) * 100
mx3_pc <- (sum(mx3_reclass[] > 0, na.rm = T) / 238627168) * 100

# 0.75 to 1.00
m <- c(0, 0.75, 0, 0.75, 1.0, 1)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
rclmat

# reclassify the raster 
mlr4_reclass <- reclassify(mlr, rclmat)
mx4_reclass <- reclassify(mx, rclmat)

# sum number of pixels above 0
# multiply by the square of the resolution to get the area
mlr4_km <- (sum(mlr4_reclass[] > 0, na.rm = T) * res(mlr4_reclass)[1]^2) / 1000 
mx4_km <- (sum(mx4_reclass[] > 0, na.rm = T) * res(mx4_reclass)[1]^2) / 1000

# get the percentage of cells
# number of cells above 0, divded by the total number of cells in the entire raster multipled by 100
mlr4_pc <- (sum(mlr4_reclass[] > 0, na.rm = T) / 238627168) * 100
mx4_pc <- (sum(mx4_reclass[] > 0, na.rm = T) / 238627168) * 100
```

Create a data frame with the suitable habitat km2 and % of suitable habitat and write out
```{r}
# make tibble
area_table <- tibble(
  thresholds = c("0 - 0.25", "0.25 - 0.5", "0.5 - 0.75", "0.75 - 1.0"),
  susceptibility = c("low", "medium", "high", "very high"),
  LR_percent = c(mlr1_pc, mlr2_pc, mlr3_pc, mlr4_pc),
  LR_km = c(mlr1_km, mlr2_km, mlr3_km, mlr4_km),
  Maxent_percent = c(mx1_pc, mx2_pc, mx3_pc, mx4_pc),
  Maxent_km = c(mx1_lm, mx2_km, mx3_km, mx4_km)
) %>% 
  mutate(LR_percent = round(LR_percent, 2),
         Maxent_percent = round(Maxent_percent, 2))

# write out csv
write_csv(area_table, paste(table_path, "/area_table.csv", sep = ""))
```
